From Real-time Attention Assessment to ``With-me-ness'' in Human-Robot Interaction

### 2015-09-22 paper.tex
Working on conclusion

In this work we have presented an on-line vision-based engagement assessment
solution that can be extended to work in different platforms for simulating
complex interaction scenarios monitoring several VFoA with a maker-less
technology. It provides some benefits like the decrement of the system
complexity as well as the increment of the naturalness of the interaction due to
the use of an RGB sensor.

@@

Gaze information acquired during the interactions is not only useful to detect
and predict engagement, but also allows the possibility to develop actively the
concept gaze-aware robot. In other words, make the system able to estimate where
the subject looks at, but also imitate the same subject's VFoA when the activity
state allows it.

%ACKNOWLEDGMENTS are optional
\section*{Acknowledgments}

>>>>

As hinted by the title of this article, we have presented here how a robot can
effectively assess in real-time and with a regular camera the focus of attention
of its interactant, and how we can, in turn, combine it with the robot's {\it a priori}
about the interaction to build a straight-forward (yet novel in HRI) metric of
\emph{with-me-ness} over the course of the interaction.

\paragraph{Practicality and Relevance of 2D Methods for Head Pose Estimation}

@@

%Gaze information acquired during the interactions is not only useful to detect
%and predict engagement, but also allows the possibility to develop actively the
%concept gaze-aware robot. In other words, make the system able to estimate where
%the subject looks at, but also imitate the same subject's VFoA when the activity
%state allows it.

\paragraph{With-me-ness: yet another metric of engagement?}

Borrowing the neologism from the field of CSCL, we have also introduced here
\emph{with-me-ness} as a measure of ``how much the user in \emph{with} the robot during
a task''. Interestingly, this can be measured over the course the interaction,
thus providing the robot with a relatively high-level social construct,
undoubtedly related to engagement.

One may reasonably wonder how different is \emph{with-me-ness} from \emph{joint
attention} on one hand, and from \emph{engagement} on the other hand.
With-me-ness is related to both, with however noteworthy nuances: (Triadic) \emph{joint
attention} is understood as the cognitive realization of a shared attention to
an object, itself building on a shared perception of that object (\ie joint attention
builds on a \emph{perceptual} alignment of two agents). \emph{Conceptual
with-me-ness} as proposed by Sharma~\etal is on the other hand \emph{referential}: ``you are
\emph{with} me if you focus on what I refer to, either explicitly or
implicitly''. We propose to understand it here in a slightly broader way, that
better reflect the \emph{interaction}: ``you are \emph{with} me if you focus on
what is important for the interactive task at hand.''

On the other hand, with-me-ness is only a precursor of engagement: it
does not say much about the cognitive commitment of a user to an interaction. A
user may closely adhere to the injunctions of the robot (or, actually, of the
experimenters) without being \emph{engaged} into the interaction. This is
typically seen in child-robot interaction: children will attempt to closely
follow what they are asked to do -- which may \emph{look like} they are engaged
into the interaction -- while they merely \emph{obey orders}.

One of the strength of with-me-ness is its objectivity: we can formalize it, and
as such, it is valuable to assess and compare how users are willing or able to
interact with a robot. Besides, we have hopefully demonstrated in this article
that with-me-ness is an operational \emph{in-the-moment} metric that can be also
used as feed-back to the robot controller to build richer, more adaptive
interactive behaviors for our robots.

\section*{Acknowledgments}
